The Anatomy of a Temporal Summarization System

### 2015-05-26 experiments.tex
Addressed comments. Fixed margins/sizings. Added authors.

\textbf{Affinity Propagation only (AP) } The purpose of this model
is to directly measure the effect of integrating salience and clustering by providing a baseline that uses the identical 
clustering component, but without the salience information.
In this model, input sentences are apriori equally likely to be exemplars;
%clustering algorithm without the salience predictions.
the salience values are uniformly set as the median value of the 
input similarity scores, as is commonly used in the AP literature \cite{frey2007clustering}.
After clustering a sentence batch, the exemplars are examined in order
of increasing time since event start and selected as updates if their
maximum similarity to the previous updates is less than $\lambda_{sim}$,
as in the novelty filtering stage of AP+Salience.

\textbf{Hierarchichal Agglomerative Clustering (HAC)} : We provide another clustering baseline, single-linkage hierarchichal agglomerative clustering.
We include this baseline to show that AP+Salience is not just an improvement
over AP but centrality driven methods in general.
HAC was chosen over other clustering approaches because the number of clusters 
is not an explicit hyper-parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
using the cophenetic distance criteria, i.e. observations in each flat cluster have no greater a cophenetic distance than a threshold.
Cluster centers are determined to be the sentence with highest cosine similariy to the flat cluster mean.
Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold $\lambda_{sim}$ as is done in the AP model.

\textbf{Rank by Salience (RS)} 
We also isolate the impact of our salience model in order to demonstrate 
the that the fusion of clustering and salience prediction improves over
predicting salience alone.
In this model we predict the salience of sentences as in step 1 for 
AP+Salience. %However, rather than clustering, we simply rank the sentences
%by salience.
We omit the clustering phase (step 2).
Updates are selected identically to step 3 of AP+Salience, proceeding 
in order of decreasing salience, selecting updates that are above a salience 
threshold $\lambda_{sal}$ and below a similarity threshold $\lambda_{sim}$
with respect to the previously selected updates.

>>>>

\textbf{Affinity Propagation only (AP) } The purpose of this model
is to directly measure the effect of integrating salience and clustering by providing a baseline that uses the identical 
clustering component, but without the salience information.
In this model, input sentences are \textit{apriori} equally likely to be exemplars;
%clustering algorithm without the salience predictions.
the salience values are uniformly set as the median value of the 
input similarity scores, as is commonly used in the AP literature \cite{frey2007clustering}.
After clustering a sentence batch, the exemplars are examined in order
of increasing time since event start and selected as updates if their
maximum similarity to the previous updates is less than $\lambda_{sim}$,
as in the novelty filtering stage of AP+Salience.

\textbf{Hierarchichal Agglomerative Clustering (HAC)} We provide another clustering baseline, single-linkage hierarchichal agglomerative clustering.
We include this baseline to show that AP+Salience is not just an improvement
over AP but centrality driven methods in general.
HAC was chosen over other clustering approaches because the number of clusters 
is not an explicit hyper-parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
using the cophenetic distance criteria, i.e. observations in each flat cluster have no greater a cophenetic distance than a threshold.
Cluster centers are determined to be the sentence with highest cosine similariy to the flat cluster mean.
Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold $\lambda_{sim}$ as is done in the AP model.

\textbf{Rank by Salience (RS)} 
We also isolate the impact of our salience model in order to demonstrate 
that the fusion of clustering and salience prediction improves over
predicting salience alone.
In this model we predict the salience of sentences as in step 1 for 
AP+Salience. %However, rather than clustering, we simply rank the sentences
%by salience.
We omit the clustering phase (step 2).
Updates are selected identically to step 3 of AP+Salience, proceeding 
in order of decreasing salience, selecting updates that are above a salience 
threshold $\lambda_{sal}$ and below a similarity threshold $\lambda_{sim}$
with respect to the previously selected updates.

### 2015-05-26 introduction.tex
Addressed comments. Fixed margins/sizings. Added authors.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
We demonstrate that combining salience with clustering produces more relevant summaries
compared to baselines using clustering or relevance alone.  
Our experiments suggest that this
is because our system is better able to adapt to
dynamic changes in input volume which adversely 
affect methods that use redundancy as a proxy for salience. 
%or predict salience but
%ignore redundancy in the input alone.

>>>>

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
We demonstrate that combining salience with clustering produces more relevant summaries
compared to baselines using clustering or relevance alone.  
Our experiments suggest that this
is because our system is better able to adapt to
dynamic changes in input volume that adversely 
affect methods that use redundancy as a proxy for salience. 
%or predict salience but
%ignore redundancy in the input alone.

### 2015-05-26 methods.tex
Addressed comments. Fixed margins/sizings. Added authors.

In order to measure the usefulness of a system's updates, we consider
the degree to which the system output reflects the different
aspects of an event.  Events are often composed of a variety of sub-events.  
For example, the Hurricane Sandy
event includes sub-events related to the storm making landfall,
the ensuing flooding, the many transportation issues, amongst many
others.  An ideal system would update the user about each of these
sub-events as they occur, with low latency.  
%In order to use
%language consistent with previous literature, 
We refer to these
sub-events as the \emph{nuggets} associated with an event.  A nugget is
defined as a fine-grained atomic sub-event associated with an event.  
%An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
We present 2 example nuggets associated with the Hurricane
Sandy event 
in Figure \ref{fig:nuggets}. Note that each event has anywhere from 50 to
several hundred nuggets in total.
We describe how these 
nuggets are found in Section \ref{sec:data}.

@@

\textbf{Basic Features}
We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.  
The data stream comprises text extracted from raw html documents;
these features help to downweight sentences that are not 
content (e.g. web page titles, links to other content) or
more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

\textbf{Query Features}
Query features measure the relationship between the sentence and the event query and type.  These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.  
For example, for event type \emph{earthquake},  we match sentence terms 
``quake'', ``temblor'', ``seism'', and ``aftershock''.

@@

\textbf{Geographic Relevance Features}
%\fdcomment{need introduction here.  what are geographic features?}
The disasters in our corpus are all phenomena that
affect some part of the word. 
Where possible, we would like to capture a sentence's proximity to the event,
i.e. when a sentence references a location, it should be close to the 
area of the disaster.

There are two challenges to using geographic features. First we do not 
know where the event is and second most sentences do not contain references
to a location.
We address the first issue by extracting all locations from 
documents relevant to the event at the
current hour and looking up their latitude and 
longitude using a publicly available geolocation service. 
%\fdcomment{is this all documents in the bucket or the filtered set?  is it a biased sample?}
Since the documents that are at least somewhat relevant to the event,
we assume in aggregate the locations should give us a rough area of interest.
The locations are clustered %(using affinity propagation and uniform salience)
%\fdcomment{affinity propagation undefined}
and we treat the resulting cluster centers
as the event locations for the current time.

To address the second issue, 
we compute features for the document as a whole and all
sentences from that document receive the same feature value.

Using the locations in each document, we compute the median distance to the 
nearest event location. Because document position is a good indicator 
of importance we also compute the distance of the first mentioned
location to the nearest event location. All sentences in the document take
as features these two distance calculations.
Because some events can move, we also
compute these distances to event locations from the previous hour.

@@

\subsubsection{Model}

At run-time
our summarization system
does not have access to event nuggets to compute sentence salience. 
However, we can use the nuggets
in our training data to model salience as a function of our features.
Such a model generalizes to new events.
For each event in our training data, we sample a set of sentences and  each 
sentence's salience is computed according to Equation \ref{eq:salience}.
This results in a training set of sentences and the target values 
(their salience) to predict.

Given a feature representation for the sentences, we can use
any regression technique to predict salience.
We opt to use a Gaussian process (GP) regression model
\cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
kernel.  
Our features fall naturally into five groups and we use a separate RBF kernel
for each, using the sum of each feature group RBF kernel as the final input
to the GP model.
%GP regressors are a
%class of data-driven, non-parametric model generalizing the multi-variate
%Gaussian to the infinite dimensional setting.

@@

Once we have predicted the salience for a batch of sentences, we must
now select a set of update candidates, i.e. sentences that are both salient
and representative of the current batch.
To accomplish this, we combine the output of our 
salience prediction model with the affinity
propagation algorithm.
%to identify a set of exemplar sentences 
%for each input batch. 
Affinity propagation (AP) is a clustering algorithm
that identifies a subset of datapoints as exemplars and forms clusters
by assigning the remaining points to one of the exemplars 
\cite{frey2007clustering}. AP attempts to 
maximize the net similarity objective 
\[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
+ \sum_{i : i = e_i}^n \operatorname{salience}(e_i)  \]
where $e_i$ is the exemplar of the $i$-th data point, and functions
$\operatorname{sim}$ and $\operatorname{salience}$ express the pairwise 
similarity
of data points and our predicted apriori preference of a data point to be an exemplar
respectively. 
AP differs from other $k$-centers algorithms in that it simultaneously 
considers all data points as exemplars, making it less prone to finding
local optima as a result of poor initialization. Furthermore, the 
second term in $\mathcal{S}$ incorporates the individual importance of 
data points as candidate exemplars; most other clustering algorithms only make
use of the first term, i.e. the pairwise similarities between data points.

@@

%In our summarization system we use the output of our salience model as 
%the preferences, i.e.
%$\operatorname{salience}(s) = \operatorname{pref}(s)$; for the similarity 
%function we use semantic similarity. 
Recall that $\operatorname{salience}(s)$
is a prediction of the semantic similarity of $s$ to information about the 
event be summarized, i.e. the set of event nuggets.
Intuitively, when maximizing objective function $\mathcal{S}$, AP must balance
between best representing the input data and representing the most salient
input. Additionally, when the level of input is high but the salience
predictions are low, the preference term will guide AP toward a solution
with fewer clusters; vice-versa when input is very salient on average but
the volume of input is low. The adaptive nature of our model differentiates
our method from most other update summarization systems.

>>>>

In order to measure the usefulness of a system's updates, we consider
the degree to which the system output reflects the different
aspects of an event.  Events are often composed of a variety of sub-events.  
For example, the Hurricane Sandy
event includes sub-events related to the storm making landfall,
the ensuing flooding, the many transportation issues, among many
others.  An ideal system would update the user about each of these
sub-events as they occur. %, with low latency.  
%In order to use
%language consistent with previous literature, 
We refer to these
sub-events as the \emph{nuggets} associated with an event.  A nugget is
defined as a fine-grained atomic sub-event associated with an event.  
%An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
We present 2 example nuggets associated with the Hurricane
Sandy event 
in Figure \ref{fig:nuggets}. 
Note that each event has anywhere from 50 to
several hundred nuggets in total in our gold dataset.
%Note that each event has anywhere from 50 to
%several hundred nuggets in total.
We describe how these 
nuggets are found in Section \ref{sec:data}.

@@

\textbf{Basic Features}
We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, and number of named entities.  
The data stream comprises text extracted from raw html documents;
these features help to downweight sentences that are not 
content (e.g. web page titles, links to other content) or
more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

\textbf{Query Features}
Query features measure the relationship between the sentence and the event 
query and type.  These include the number of query words present in the 
sentence in addition to the number of event type synonyms, hypernyms, and 
hyponyms using WordNet \cite{miller1995wordnet}.  
For example, for event type \emph{earthquake},  we match sentence terms 
``quake'', ``temblor'', ``seism'', and ``aftershock''.

@@

\textbf{Geographic Relevance Features}
%\fdcomment{need introduction here.  what are geographic features?}
The disasters in our corpus are all phenomena that
affect some part of the world. 
Where possible, we would like to capture a sentence's proximity to the event,
i.e. when a sentence references a location, it should be close to the 
area of the disaster.

There are two challenges to using geographic features. First we do not 
know where the event is, and second, most sentences do not contain references
to a location.
We address the first issue by extracting all locations from 
documents relevant to the event at the
current hour and looking up their latitude and 
longitude using a publicly available geolocation service. 
%\fdcomment{is this all documents in the bucket or the filtered set?  is it a biased sample?}
Since the documents that are at least somewhat relevant to the event,
we assume in aggregate the locations should give us a rough area of interest.
The locations are clustered %(using affinity propagation and uniform salience)
%\fdcomment{affinity propagation undefined}
and we treat the resulting cluster centers
as the event locations for the current time.

%\ckcomment{(Paragraphs in blue should replace last two paragraphs in geo relevance section)\\
The second issue arises from the fact that the majority of sentences in our 
data do not contain explict references to locations, i.e. a sequence of tokens
tagged as location named entities.   
Our intuition is that geographic relevance is important in the disaster
domain, and we would like to take advantage of the sentences that do have 
location information present. To make up for this imbalance, we  
%Extracting features for just these sentences
%would leave 
% Last sentence is unclear
%Assigning a null or default value to sentences without explicit location is 
%not a viable option as 
%the imbalance of null/default values would make it difficult for the model
%to learn meaningful weights for these features.
% also unclear
instead compute an overall location for the document and derive 
geographic features based on the document's proximity to the event in 
question.
These features are assigned to all sentences in the document.\\\\
Our method of computing document-level geographic relevance features is as 
follows.
Using the locations in each document, we compute the median distance to the 
nearest event location. Because document position is a good indicator 
of importance we also compute the distance of the first mentioned
location to the nearest event location. All sentences in the document take
as features these two distance calculations.
Because some events can move, we also
compute these distances to event locations from the previous hour.
%}

%To address the second issue, 
%we compute features for the document as a whole and all
%sentences from that document receive the same feature value.

%Using the locations in each document, we compute the median distance to the 
%nearest event location. Because document position is a good indicator 
%of importance we also compute the distance of the first mentioned
%location to the nearest event location. All sentences in the document take
%as features these two distance calculations.
%Because some events can move, we also
%compute these distances to event locations from the previous hour.

@@

\subsubsection{Model}
%\ckcomment{(replace section with this)\\
Given our feature representation of the input sentences, we need only target 
salience values for model learning. 
For each event in our training data, we sample a set of sentences and  each 
sentence's salience is computed according to Equation \ref{eq:salience}.
This results in a training set of sentences, their feature representations,
and their target salience values to predict.

%Given the feature representation/salience pairs, we can use
%any regression technique to predict salience.
We opt to use a Gaussian process (GP) regression model
\cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
kernel for the salience prediction task.  
Our features fall naturally into five groups and we use a separate RBF kernel
for each, using the sum of each feature group RBF kernel as the final input
to the GP model.
%}

%At run-time
%our summarization system
%does not have access to event nuggets to compute sentence salience. 
%However, we can use the nuggets
%in our training data to model salience as a function of our features.
%Such a model generalizes to new events.
%For each event in our training data, we sample a set of sentences and  each 
%sentence's salience is computed according to Equation \ref{eq:salience}.
%This results in a training set of sentences and the target values 
%(their salience) to predict.
%
%Given a feature representation for the sentences, we can use
%any regression technique to predict salience.
%We opt to use a Gaussian process (GP) regression model
%\cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
%kernel.  
%Our features fall naturally into five groups and we use a separate RBF kernel
%for each, using the sum of each feature group RBF kernel as the final input
%to the GP model.

%GP regressors are a
%class of data-driven, non-parametric model generalizing the multi-variate
%Gaussian to the infinite dimensional setting.

@@

Once we have predicted the salience for a batch of sentences, we must
now select a set of update candidates, i.e. sentences that are both salient
and representative of the current batch.
To accomplish this, we combine the output of our 
salience prediction model with the affinity
propagation algorithm.
%to identify a set of exemplar sentences 
%for each input batch. 
Affinity propagation (AP) is a clustering algorithm
that identifies a subset of datapoints as exemplars and forms clusters
by assigning the remaining points to one of the exemplars 
\cite{frey2007clustering}. AP attempts to 
maximize the net similarity objective 
\[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
+ \sum_{i : i = e_i}^n \operatorname{salience}(e_i)  \]
where $e_i$ is the exemplar of the $i$-th data point, and functions
$\operatorname{sim}$ and $\operatorname{salience}$ express the pairwise 
similarity
of data points and our predicted \textit{apriori} preference of a data point to be an exemplar
respectively. 
AP differs from other $k$-centers algorithms in that it simultaneously 
considers all data points as exemplars, making it less prone to finding
local optima as a result of poor initialization. Furthermore, the 
second term in $\mathcal{S}$ incorporates the individual importance of 
data points as candidate exemplars; most other clustering algorithms only make
use of the first term, i.e. the pairwise similarities between data points.

@@

%In our summarization system we use the output of our salience model as 
%the preferences, i.e.
%$\operatorname{salience}(s) = \operatorname{pref}(s)$; for the similarity 
%function we use semantic similarity. 
Recall that $\operatorname{salience}(s)$
is a prediction of the semantic similarity of $s$ to information about the 
event be summarized, i.e. the set of event nuggets.
Intuitively, when maximizing objective function $\mathcal{S}$, AP must balance
between best representing the input data and representing the most salient
input. Additionally, when the level of input is high but the salience
predictions are low, the preference term will guide AP toward a solution
with fewer clusters; \textit{vice-versa} when input is very salient on average but
the volume of input is low. The adaptive nature of our model differentiates
our method from most other update summarization systems.

### 2015-05-26 results.tex
almost final draft
Addressed comments. Fixed margins/sizings. Added authors.

\begin{figure}
    \includegraphics[]{rouge-time.eps}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\vspace{-14pt}
\end{figure}

@@

\subsection{Expected Gain and Comprehensiveness}
\begin{figure}
  \includegraphics[]{nuggets-metrics.eps}
\caption{Expected Gain and Comprehensiveness performance.}
\label{fig:nperf}
\vspace{-12pt}
\end{figure}

@@

When looking at the comprehensiveness of the summaries AP outperforms
 AP+Salience. The compromise encoded in the AP+Salience objective
function, between being representitive and being salient, is seen clearly here
where the performance of the AP+Salience methods is lower bounded by 
the salience focused RS
system and upper bounded by the clustering only AP system.
Overall, AP+Salience achieves the best balance of these two metrics.

@@

Table~\ref{tab:farouge} shows the results of our feature ablation
tests. Removing the language models yields a statistically 
significant drop in both ngram recall and F-measure. 
Interestingly, removing the basic features leads to an
increase in both unigram and bigram precision; in the bigram
case this is enough to cause a statistically significant increase
in F-measure over the full model. In other words, the generic features
actually lead to an inferior model when we can incorporate more appropriate
domain specific features.
%This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.

>>>>

\begin{figure}
    \includegraphics[]{rouge-time.eps}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
%\vspace{-14pt}
\end{figure}

@@

\subsection{Expected Gain and Comprehensiveness}
\begin{figure}[h]
  \includegraphics[]{nuggets-metrics.eps}
\caption{Expected Gain and Comprehensiveness performance.}
\label{fig:nperf}
%\vspace{-12pt}
\end{figure}

@@

When looking at the comprehensiveness of the summaries AP outperforms
 AP+Salience. The compromise encoded in the AP+Salience objective
function, between being representative and being salient, is seen clearly here
where the performance of the AP+Salience methods is lower bounded by 
the salience focused RS
system and upper bounded by the clustering only AP system.
Overall, AP+Salience achieves the best balance of these two metrics.

@@

Table~\ref{tab:farouge} shows the results of our feature ablation
tests. Removing the language models yields a statistically 
significant drop in both ngram recall and F-measure. 
Interestingly, removing the basic features leads to an
increase in both unigram and bigram precision; in the bigram
case this is enough to cause a statistically significant increase
in F-measure over the full model. In other words, the generic features
actually lead to an inferior model when we can incorporate more appropriate
domain specific features.
The result mirrors Sparck Jones' claim that generic approaches to summarization cannot produce a useful summary \cite{ksj98}.
%This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.
